---
title: "Tidyverse"
output: html_document
editor_options: 
  chunk_output_type: console
---

### Install packages 
```{r eval=FALSE}
install.packages("here")
install.packages("tidyverse")
install.packages("magrittr")
install.packages("dplyr")
install.packages("broom")
```

### Load libraries & data
Load libraries
```{r load-packages,  message=FALSE, warning=FALSE}
library(here) # Look where your folder begins.
library(dplyr)
library(magrittr)
library(tidyverse)
library(broom)
```

### Save this built-in dataset as a dataframe
The "The Effect of Vitamin C on Tooth Growth in Guinea Pigs" dataset contains data from an experiment wherein 60 guinea pigs were given vitamin C either as pure vitamin C (VC) or orange juice (OJ) (column "supp") in doses of 0.5, 1, or 2 mg/day (column "dose"). The researchers recorded the growth of their teeth (column "len").
```{r}
teeth <- ToothGrowth
```

<br/>

# The tidyverse
## What is it?
[The tidyverse "is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures."](https://www.tidyverse.org/) The tidyverse contains some extremely useful packages such as

- <code>[**ggplot2**](https://ggplot2.tidyverse.org/)</code>, which lets you make fantastic graphics;
- <code>[**tidyr**](https://tidyr.tidyverse.org/)</code>, which helps you "tidy" (clean) up your data; and
- <code>[**dplyr**](https://dplyr.tidyverse.org/)</code>, which lets you manipulate your data, like you would with pliers (hence d-*plyr*).
<br/>

***
# Exercises

#### Today we will work through examples of these tidyverse functions:

- **<code>mutate()</code>**
- **<code>group_by()</code>**
- **<code>summarize()</code>**
- **<code>filter()</code>**
- **<code>select()</code>**

#### Then we'll introduce some functions from the **<code>broom</code>** package that are especially useful for working with linear models such as 

- **<code>tidy()</code>**
- **<code>augment()</code>**
- **<code>glance()</code>**


<br/>
  
#### Prequel: **<code>magrittr</code>**'s pipes
A fantastic tool that is worth getting to know is the **<code>magrittr</code>** package's collection of operators called *pipes*. As stated on their [webpage](https://magrittr.tidyverse.org/),

> "The **<code>magrittr</code>** package offers a set of operators which make your code more readable by:

> - structuring sequences of data operations left-to-right (as opposed to from the inside and out),
> - avoiding nested function calls,
> - minimizing the need for local variables and function definitions, and
> - making it easy to add steps anywhere in the sequence of operations."  
  
##### We'll frequently set up our code like this:

<code>new_object <- old_object %>% function()</code>


##### If you'd rather overwrite your original object, you can use either of these setups:

<code>old_object <- old_object %>% function()</code>

<code>old_object %<>% function()</code> *(notice that the pipe has two arrows inside it now)* <br/><br/>

***

## **<code>mutate()</code>**
The <code>mutate()</code> function lets us create new columns based on existing columns. In this example, we'll create a new column that takes the "len" data- guinea pig tooth length measured in mm- and divide it by 100 to create a similar column with cm as the unit of length.

The . after mutate( refers to how we're performing the <code>mutate()</code> function on the object already named after the assignment arrow (the teeth object).
```{r}
teeth <- teeth %>%
  dplyr::mutate(.,
                len_cm = len / 100)

head(teeth)
```

We can also mutate multiple columns at once. Let's add a new column that calculates the mean tooth length for all guinea pigs:
```{r}
teeth <- teeth %>%
  dplyr::mutate(.,
                len_cm = len / 100,
                mean_len = mean(len))

head(teeth)
```

<br/>

## **<code>group_by()</code>**
So now we know the mean tooth length for all guinea pigs, but that's not very helpful. It'd be more informative to know the mean tooth length for specific groups of guinea pigs- based on the doses of vitamin C adminstered and the method of administration.
First, let's group the subjects by the method of administration (supp) to find the mean tooth length per group (VC or OJ):
```{r}
teeth_2 <- teeth %>%
  dplyr::group_by(supp) %>%
  dplyr::mutate(.,
         mean_len = mean(len))

head(as.data.frame(teeth_2), 60)
```


Now let's incorporate dosage (dose) as another grouping so that the subjects are grouped by both supp *and* dose:
```{r}
teeth_2 <- teeth %>%
  dplyr::group_by(supp, dose) %>%
  dplyr::mutate(.,
         mean_len = mean(len))

head(as.data.frame(teeth_2), 60)
```

That's more like it!

However, there's a better way to organize this data...

<br/>

## **<code>summarize()</code>**
If you don't need to keep the individual subjects' tooth length data, then it makes sense to use <code>summarize()</code> to condense our summarized data into fewer rows. Our output file will have one row for each combination of supp x dose.
```{r}
teeth_3 <- teeth %>%
  dplyr::group_by(supp, dose) %>%
  dplyr::summarize(mean_len = mean(len))

head(teeth_3)
```

You can even summarize multiple times, such as if you wanted to first get mean values per supp x dose group and then average *those* values to get mean length per supp group.
```{r}
teeth_4 <- teeth %>%
  dplyr::group_by(supp, dose) %>%
  dplyr::summarize(mean_len = mean(len)) %>%
  dplyr::group_by(supp) %>%
  dplyr::summarize(mean_len = mean(mean_len))

head(teeth_4)
```

What if you just want to summarize all the rows of your dataset? Then you can use <code>summarize()</code> in conjunction with the <code>across()</code> "helper" function.


Let's calculate the mean values for dose and len_cm. To do so, you'd set up your code like this. Be sure you're not selecting any non-numeric columns, like supp.
```{r}
teeth_5 <- teeth %>%
  dplyr::summarize(across(dose:len_cm, mean))

teeth_5
```


You can even summarize multiple variables in multiple ways, like in this example, where we're calculating both mean, maximum, and minimum values for dose and length. In the set of parentheses determining which calculations to perform (mean, max, min), be sure to put the suffix you want to append to the column name first (e.g. mean, maximum, minimum), then the = sign, then the R-recognized calculation name (mean, max, min).
```{r}
teeth_6 <- teeth %>%
  dplyr::summarize(across(c(len, dose),
                          c(mean = mean,
                            maximum = max,
                            minimum = min))) %T>%
  print()
```
*You'll also notice that I added a special pipe- **%T>%**- plus the <code>print()</code> function at the end of the command. This tells R to perform the functions after the special pipe but to not associate those functions with the object being created (teeth_6). This pipe is very useful when creating plots, for instance.*


Let's imagine that you only want to focus on the guinea pigs given orange juice. How would you get rid of the other rows?

<br/>

## **<code>filter()</code>**
The <code>filter()</code> function lets you filter out the rows you don't want. Here we've added a line of code to keep only the rows where the guinea pig was given vitamin C as orange juice (OJ), not as pure vitamin C.
```{r}
teeth_7 <- teeth %>%
  dplyr::filter(supp == "OJ")

head(teeth_7)
```

<br/>

## **<code>select()</code>**
Similarly, the <code>select()</code> function lets you filter out the *columns* you don't want. Our new dataframe only includes the first three columns: from len to dose. Both of these commands accomplishes the same goal, whether you name the columns by their names or index position.
```{r}
teeth_8 <- teeth %>%
  dplyr::select(len:dose)

head(teeth_8)


teeth_8 <- teeth %>%
  dplyr::select(1:3)

head(teeth_8)
```

<br/>

***

## **Helpful Functions for Linear Models**

Let's load a new built-in dataset: iris. This dataset offers lengths and widths of sepals and petals of three species of flowers.
```{r}
iris1 <- iris
```

We're going to set up some linear models- one for each species of flower- and then use three functions from the <code>broom</code> package to summarize important information about the models which will be generated in a format that makes it simple to compare and plot models.

As stated in the [**<code>broom</code>** overview page](https://broom.tidymodels.org/), here is the role of each function:

#### **<code>tidy()</code>**: 
*"Summarizes information about model components"*

#### **<code>augment()</code>**
*"Reports information about the entire model"*

#### **<code>glance()</code>**
*"Adds informations about observations to a dataset"*


<br/>

### The long way
#### Set up individual models
```{r}
mod1 <- lm(Sepal.Length ~ Sepal.Width,
           iris1 %>% filter(Species == "setosa"))

mod2 <- lm(Sepal.Length ~ Sepal.Width,
           iris1 %>% filter(Species == "versicolor"))

mod3 <- lm(Sepal.Length ~ Sepal.Width,
           iris1 %>% filter(Species == "virginica"))
```

<br/>

### **<code>tidy()</code>**
```{r}
broom::tidy(mod1)
# broom::tidy(mod2)
# broom::tidy(mod3)
```

<br/>

### **<code>augment()</code>**
```{r}
broom::augment(mod1)
# broom::augment(mod2)
# broom::augment(mod3)
```

<br/>

### **<code>glance()</code>**
```{r}
broom::glance(mod1)
# broom::glance(mod2)
# broom::glance(mod3)
```

<br/>

### The *fastR* way to do this:
Notice that we are using the <code>group_by</code> function perform the same command on each species but in the same chunk of code.

<br/>

### **<code>tidy()</code>**
```{r}
# tidy
mod_all.tidy <- iris1 %>%
  group_by(Species) %>%
  do(fit = broom::tidy(lm(Sepal.Length ~ Sepal.Width,
              data = .))) %>%
  unnest(fit) %T>%
  print()
```

<br/>

### **<code>augment()</code>**
```{r}
mod_all.augment <- iris1 %>%
  group_by(Species) %>%
  do(fit = broom::augment(lm(Sepal.Length ~ Sepal.Width,
              data = .))) %>%
  unnest(fit) %T>%
  print()
```

<br/>

### **<code>glance()</code>**
```{r}
mod_all.glance <- iris1 %>%
  group_by(Species) %>%
  do(fit = broom::glance(lm(Sepal.Length ~ Sepal.Width,
              data = .))) %>%
  unnest(fit) %T>%
  print()
```
