---
title: "Tips & Tricks"
output:
  html_document:
    theme: sandstone
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

The purpose of this segment is to share some of our favorite tools for working with data in R. We will focus on packages that are less popular than the Greatest of All Times (e.g. [tidyverse](https://www.tidyverse.org/) and its associated [ggplot2](https://ggplot2.tidyverse.org/), [dplyr](https://dplyr.tidyverse.org/), [tidyr](https://tidyr.tidyverse.org/), etc.) though we *highly encourage* you to check these out!

## `here`: no more getting lost in file paths

> *Use this package ~~when...~~* all the time. Make it a part of your regular coding routine. It's *that* good.


As we introduced in the previous module, `here` is an excellent package that's worth getting to know because it will let you use relative as opposed to absolute pathnames. This will simplify importing and exporting files as well as sharing them with others.

### How to install
```{r eval = FALSE}
install.packages("here")
```

### How to use
```{r}
library(here)
here::here()   # once you call the package with the library(here) call, you can use this function to remind you where your project root begins.
```


<style>
.btn {
  border: 2px solid black;
  background: gray;
  background-color: gray;
  padding: 14px 28px;
  font-size: 14px;
  cursor: pointer;
  display: inline-block;
  border-color: green;
  color: green;
}
.btn:hover {background: #5DADE2;}
</style>

<a class="btn" href="https://sbreitbart.github.io/BGSS_Retreat_2021_Workshop/Functions/Functions.html"> revisit "here" package description </a>

```{r, echo=FALSE}
# Define variable containing url
url <- "https://raw.githubusercontent.com/allisonhorst/stats-illustrations/master/rstats-artwork/here.png"
```

<center><img src="`r url`"></center>
Artwork by [Allison Horst](https://github.com/allisonhorst). `set_wd`, be gone!





## `magrittr`: these pipes will make your work *flow*

> *Use this package when...* 


### How to install
```{r eval = FALSE}
install.packages("magrittr")
```

### How to use
```{r}
library(magrittr)
```




## `patchwork`: make your figures nice and cozy

> *Use this package when...* 


### How to install
```{r eval = FALSE}
install.packages("patchwork")
```

### How to use
```{r}
library(patchwork)
```





## `performance`: evaluate your general linear models in a flash

> *Use this package when...* 


### How to install
```{r eval = FALSE}
install.packages("performance")
```

### How to use
```{r}
library(performance)
```






## `annotater`: *finally* remember why you loaded all those packages!

> *Use this package when...* you forget why you loaded packages at the top of your R script/notebook/markdown file OR you want to clarify why you did so for collaborators. (+1 points for reproducible science!)

It happens: you start your R file with a list of packages to be loaded with your `library()` call. You constantly add to it, listing more packages who functions you use to complete your analysis. Over time, you figure out that your advisors would prefer if you didn't use the [Wes Anderson color palette](https://github.com/karthik/wesanderson), or that you're better off creating figures with `patchwork` vs `cowplot` (sorry, `cowplot`). So, if you're anything like me, after learning stats *and* R simultaneously while doing your first thesis project, you end up with a very impressive list of R packages, half of which you can't remember why you loaded in the first place.

Have no fear! This is where the genius of the [`annotater`](https://github.com/luisDVA/annotater) package comes in to save us (and those who try to read our code, bless them)!



### How to install
```{r eval = FALSE}
install.packages("remotes")
remotes::install_github("luisDVA/annotater")
```

### How to use
After you've installed `remotes` and `annotater`, save your R files, close RStudio, and reopen it.

Click anywhere in the Source pane (aka the one with your R files).

Navigate your cursor to the "Addins" button in the bar below the File-Edit-Code-View etc. bar. Click it and select, "Annotate package calls in active file". Voila!

```{r, echo=FALSE}
# Define variable containing url
url2 <- "https://raw.githubusercontent.com/luisdva/annotater/master/inst/media/annotcalls.gif"
```

<center><img src="`r url2`"></center>



## `dplyr`: quick manipulations of your data

> *Use this package when...* you are trying to quickly manipulate, summarize, or combine data. 

Often we are required to summarize data by certain categories, groups, or treatments. In other instances, we are looking to create a new column that contains a metric specific for our analysis. The `dplyr` package uses a pipe-format that allows for an easy creation of a workflow. Additionally, `dplyr` allows for some basic data management including selecting certain columns, renaming them, or sorting data. All of these functions are based off an `SQL` backend, which provides some familiarity for those with a data or computer science background.  


### How to install
```{r eval = FALSE}
install.packages("dplyr")
```

### How to use
```{r warning=F, message=F}
library(dplyr)

## Load sample dataset
data(starwars)

###  Calculate summary statistics for a group within the dataset. 
summmarizedData <- starwars %>% 
                      group_by(homeworld) %>%  ## select variables to summarize by
                      summarize(avgHeight = mean(height), nObs= length(height)) ## select which columns to be summarized and by which function
summmarizedData


###  Calculate summary statistics for multiple groups within the dataset. 
summmarizedData <- starwars %>% 
                      group_by(homeworld, species) %>%  ## select variables to summarize by
                      summarize(avgHeight = mean(height), nObs= length(height)) ## select which columns to be summarized and by which function
summmarizedData

## Create a new column of estimated BMI for each person
starwarsBMI <- starwars %>% 
                  mutate(bmi = mass / (height/100)^2)
starwarsBMI %>% select(name, BMI = bmi) ## show calculated data and rename it to capitalize BMI

## Find the individuals with the lowest BMI per planet
lowBMI <- starwarsBMI %>% 
            group_by(homeworld) %>% 
            slice(which.min(bmi))
lowBMI
```

A cheatsheet for frequently used functions from `dplyr` can be found [here](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)

## `tidyr`: quick manipulations of your data (cont)

> *Use this package when...* you are trying to quickly manipulate, summarize, or combine data. 

The `tidyr` package comes from the same author as `dplyr` and share the same common syntax of pipes and SQL-style structure. There are many functions found within this package but there are two that I think are extremely useful: 1) converting data between long and wide formats and 2) separating a column into multiple. 

### How to install
```{r eval = FALSE}
install.packages("tidyr")
```

### How to use
```{r warning=F, message=F}
library(tidyr)

## Load sample dataset
data(starwars)

### Select a matrix that has the individual and species. Then convert it from long to wide format
longMat <- starwars %>% 
            select(name, species, mass) %>% 
            spread(species, mass)

## convert back to wide format
wideMat <- longMat %>% 
            gather(species, mass, 2:39)


## Split colour into multiple columns
starwars %>% 
  separate(skin_color, sep=", ", into=c("MainColour","SecondaryColour","AncillaryColour")) %>% 
  select(name, MainColour, SecondaryColour, AncillaryColour)
```


## Survey

Please take 5 minutes and give us some feedback about today's workshop- we'd be so grateful! Check out the "Feedback Page" link below. Thanks!


<style>
.btn {
  border: none;
  background-color: inherit;
  padding: 14px 28px;
  font-size: 16px;
  cursor: pointer;
  display: inline-block;
  color: orange;
}
.btn:hover {background: #5DADE2;}
</style>


<a class="btn" href="https://sbreitbart/BGSS_Retreat_2021_Workshop/"> Home </a>
<a class="btn" href="https://easyretro.io/publicboard/gEdRrWVYh9ZQztnmdn94aX87HO83/b09d9a47-fe9c-4185-936d-445a0ad92428"> Feedback Page </a>